---
title: "M06-Reflection Essay-Advanced Data Wrangling"
author: "Vianey Villegas"
date: 'March 15, 2026'
format: 
  html: 
    toc: true
    toc-depth: 4
    number-sections: true
    code-link: true
    embed-resources: true
editor: visual
execute: 
  freeze: auto
---

# Introduction

This will be a reflection report based on the concepts I learned in module 06.

# Tidyverse Tricks Learned in Step 1

Describe the two processes. What are the big differences between the two? How do the differences result in the differences in final sample size? 

## New Skills Learned 

- Advanced **dplyr** data manipulation
- More complex **tidyr** data cleaning functions
- Efficient ways to transform datasets

:::callout-tip
Tidyverse helps the user transform the data to make it look cleaner and readable.
:::

# David Robinson's Screencast

What do you think about David Robinson's impromptu screencast performance in Step 2? 
 
 David Robinson's Screencast showed advanced data wrangling that was fast. 

## Why I was able to understand

 - He is clear and experienced 
 - He demonstrated a strong understanding when using tidyverse tools,

:::callout-tip
 Watching an experienced programmer will help learning how to code because they are more clear.
:::

# Initial Data Cleaning

What's your impression of the Revealjs presentation? Describe its capabilities as you learned from the video. What are its strengths and weaknesses compared with PPT?

 In Step 3, I saw **initial data cleaning** and **further data cleaning**.

## Initial Data Cleaning

 Initial data cleaning helps organize the dataset.

 - Removes unnecessary columns
 - Converts data types
 - Handles missing values 

## Further Data Cleaning
Further data cleaning focuses on deeper transformations. 

 - Filters observations 
 - creates new variables 
 - joins additional data 

## Difference between the both

 | Initial Cleaning | Further Cleaning |
|------|------|
| Basic formatting | Advanced transformation |
| Removing errors | Adding new insights |
| Preparing structure | Creating analysis-ready dataset |

# Revealjs Presentation

Briefly describe what the presentation is about. What did you learn from the experience of the Revealjs presentation? 

Revealjs is a presentation framework in Quarto that lets users creat **interactive presentations using code**.

## Capabilities 

- Slide transformation
- Embedded codes
- Embedded outputs 
- Quarto document integration 

:::panel-tabset

## Revealjs Strengths

- Interactive slides
- Easy integration with code
- Works well for technical presentations

## PPT Strengths

- Familiar interface
- Easier for non-technical users
- More visual design tools

:::

# Advanced Wrangling Concepts 

Pick three of them, describe when to use them, and give one coding example along with an output for each function. 

In step 6, I learned about relational data operations (**joining datasets**).

## 1. left_join()

The `left_join()` function merges two datasets while keeping all rows from the first dataset.

```{r}
library(dplyr)

data1 <- data.frame(id = c(1,2,3),
                    name = c("A","B","C"))

data2 <- data.frame(id = c(1,2,4),
                    score = c(90,85,88))

left_join(data1, data2, by="id")
```

### Output

| id | name | score |
|----|------|------|
|1|A|90|
|2|B|85|
|3|C|NA|

## 2. inner_join()

The `inner_join()` function returns only rows that match in both datasets.

```{r}
inner_join(data1, data2, by="id")
```

### Output

| id | name | score |
|----|------|------|
|1|A|90|
|2|B|85|

---

## 3. mutate()

The `mutate()` function creates new variables.

```{r}
df <- data.frame(score=c(70,85,90))
df %>%
  mutate(grade = ifelse(score >= 80, "Pass", "Fail"))
```

### Output

| score | grade |
|------|------|
|70|Fail|
|85|Pass|
|90|Pass|

---

# Reflection on Data Wrangling 

What did you like the most about working with the tools you learned in this module?

## What I learned:

I really liked working with **Tidyverse tools** because it really helps facilitate the reading od datasets.
 
 Advantages: 

- Clearing syntax
- Consistent structure
- Powerful transformation tools 

# Challenges in Learning Data Wrangling 

What seems to be the challenge, if any, for you when you try to master the wrangling tools? How may you be able to overcome the challenge?

## Challenged

- Remebering all the codes 
- LEARNING 
- Understanding how to relate the datasets 

It takes practice to learn how to code, but watching experienced programmers and being consistent can help overcome these challenged. 

# Conclusion

Module 06 I think is one of the most important because when your learning coding it's important to also know how to make your work readable. This module helps it look cleaner. 


